Ligeti, Stockhausen and Koenig:\\Aesthetics of Electronic Music

### 2015-09-27 manuscript.tex
Minor English improvements

More generally, understanding electronic music as \textbf{a way of thinking} rather than a way to create sound using electronic devices may shed a light over instrumental music by composers with former electronic music education as Ligeti, and even more important, on instrumental music that live side by side with electronically created music for almost 70 years.

@@

Darmstadt of post Wold War II became an important center for modern music thereby an attractive place for young composers.
The city is located in the state of Hessen in central Germany, in the American zone of occupation of those days.
The summer courses were financial supported by the US military government with two goals:
first, to propagate American values in effort to reeducate the German population in preparation for the establishment of democratic institutions;
and second, to provide a meeting place where musicians from the former fascist or fascist-occupied areas might further their musical reeducation through exposure to styles and techniques that had been prohibited during the fascist years (\cite{darmstadt_oxford}).
Since the first Darmstadt summer course, in 1948, the focus was on current musical concepts;
at the time it was, mainly, electronic and serail music\footnote{Internationales Musicinstitut Darmstadt: www.internationales-musikinstitut.de/en/summer-course/history/210-history.html}.

In addition to Darmstadt summer courses, the electronic music studios of Cologne were another prominent center for the Darmstadt-Cologne avant guard scene.
The studios were founded in 1951, focused on the creation of pure electronic or ``electroacoustic'' music.
It is believed that the term ``electroacoustic'' was coined by Meyer-Eppler, one of the founders of the studios, to distinguish themselves from musique-concr{\`e}te:
a technique that characterized the already established style of Parisian electronic music (\cite{nwdr}).

@@

Some of the hereby presented concepts are my summaries to the ideas found in the written materials, whereas others are quoted directly from the source.
For those, I add clarification only when absolutely necessary, as I think that most of the concepts are relatively self-explanatory.

@@

The last category is the one for higher level expressions, and is occupied almost solely by Ligetis' ideas.
These concepts based on expressions and feelings, and are characterized by descriptions that are not musical in their nature.

@@

Beyond the general concepts above, this ideas are perhaps the most dominants in this composition.
Examples can be found in the 0:00 - 1:55, 4:20 - 4:35, 5:28 - 6:25.

@@

Although this idea can be understood outside of serial thinking, as I argued before, I'm unable to find its occurrence in Ramifications.
This concept is, perhaps, too contradicted to Ligeti's way of thinking.
I will refer to this open issue again later, in the discussion.

@@

This idea shines throughout Ramifications.
The opening of the piece serve as a good example.
From the start, Ligeti `represent' a very narrow pitch range, using a sound plane.
This sound plane is in a transitory mode most of the time;
the sound repeatedly moved back and forth between dense textures and recognizable melodic lines.
In addition, after one minute the plane starts to move and expand its higher bound to higher pitches.
This trend accelerate till 2:19, where, by slowing the patterns and reducing the layer, the plane transitions into a unison.

>>>>

More generally, understanding electronic music as \textbf{a way of thinking} rather than a way to create sound using electronic devices may shed a light over instrumental music by composers with former electronic music education as Ligeti, and even more important, on instrumental music that lives side by side with electronically created music for almost 70 years.

@@

Darmstadt of post Wold War II became an important center for modern music and thereby an attractive place for young composers.
The city is located in the state of Hessen in central Germany, in the American zone of occupation of those days.
The summer courses were financial supported by the US military government with two goals:
first, to propagate American values in effort to reeducate the German population in preparation for the establishment of democratic institutions;
and second, to provide a meeting place where musicians from the former fascist or fascist-occupied areas might further their musical reeducation through exposure to styles and techniques that had been prohibited during the fascist years (\cite{darmstadt_oxford}).
Since the first Darmstadt summer course, in 1948, the focus was on current musical concepts;
at the time it was, mainly, electronic and serail music\footnote{Internationales Musicinstitut Darmstadt: www.internationales-musikinstitut.de/en/summer-course/history/210-history.html}.

In addition to Darmstadt summer courses, the electronic music studios of Cologne --- the NWDR --- were another prominent center for the Darmstadt-Cologne avant guard scene.
The studios were founded in 1951, focused on the creation of pure electronic or ``electroacoustic'' music.
It is believed that the term ``electroacoustic'' was coined by Meyer-Eppler, one of the founders of the studios, to distinguish themselves from musique-concr{\`e}te:
a technique that characterized the already established style of Parisian electronic music\footnote{Martian Arts -- The Birth of Electronic Music Recording Studios: http://www.martianarts.net/web/culture/technology/105-the-birth-of-electronic-music-recording-studios}.

@@

Some of the following concepts are my summaries to the ideas found in the written materials, whereas others are quoted directly from the source.
For them, I add clarification only when absolutely necessary, as I think that most of the concepts are relatively self-explanatory.

@@

The last category is the one for higher level expressions, and is occupied almost solely by Ligetis' ideas.
These concepts are based on expressions and feelings, and are characterized by descriptions that are not musical in their nature.

@@

Beyond the general concepts above, this ideas are perhaps the most dominants in this composition.
Examples can be found in 0:00 - 1:55, 4:20 - 4:35, 5:28 - 6:25.

@@

Although this idea can be understood outside of serial thinking, as I argued before, I'm unable to find its occurrence in Ramifications.
This concept is, perhaps, too contradicted to Ligeti's way of thinking.
I will refer to this issue again later, in the discussion.

@@

This idea shines throughout Ramifications.
The opening of the piece serves as a good example.
From the start, Ligeti `represent' a very narrow pitch range, using a sound plane.
This sound plane is in a transitory mode most of the time;
the sound repeatedly moved back and forth between dense textures and recognizable melodic lines.
In addition, after one minute the plane starts to move and expand its higher bound to higher pitches.
This trend accelerate till 2:19, where, by slowing the patterns and reducing the layers, the plane transitions into an unison.
